Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}

### 2017-09-29 Paper.tex
Testing to fix passing build, for comparison

%\usepackage{longtable}
% https://stackoverflow.com/a/1918465/7438857
%tabularx seems better, but we can include this if we need tables to break, as well as tabu and longtabu

@@

\usepackage{tabu}

@@

\begin{tabularx}{\textwidth}{rlrrX}
\toprule
0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}. Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$. As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. Otherwise it is exactly equivalent to {\small CALL} except: \\
&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{s}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_\mathbf{p}, 0, \hyperlink{I v}{I_\mathbf{v}}, \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathbf{v}} \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$. This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
\midrule
0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{SELFDESTRUCT}{}\\
\midrule
0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_\mathbf{a} \}$ \\
&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} + \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ \\
&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{b} \equiv 0$ \\
&&&& $A'_{r} \equiv A_{r} + \begin{cases}
\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
0 & \text{otherwise} 
\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{selfdestruct}} + \begin{cases}
\hyperlink{Gnewaccount}{G_{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
0 & \text{otherwise}
\end{cases}$ \\
\bottomrule
\end{tabularx}

@@

\section{Ethash}\label{app:ethash}
\subsection{Definitions}

We employ the following definitions:

%\begin{tabularx}{\textwidth}{lrX}
\begin{tabu}{X[,l]X[-r]X[-l]} 
\toprule
Name & Value & Description \\
\midrule 
\hypertarget{Jwordbytes}{}
$J_{\rm{wordbytes}}$ & 4  & This is the bytes in a word. \hypertarget{Jdatasetinit}{ \\
$J_{\rm{{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
$J_{\rm{{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
$J_{\rm{{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis. }\hypertarget{Jcachegrowth}{\\
$J_{\rm{{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch. }\hypertarget{Jepoch}{\\
$J_{\rm{{epoch}}$ & 30000 & This is the blocks per epoch. }\hypertarget{Jmixbytes}{\\
$J_{\rm{{mixbytes}}$ & 128 & This is the mix length in bytes. }\hypertarget{Jhashbytes}{\\
$J_{\rm{{hashbytes}}$ & 64 & This is the hash length in bytes. }\hypertarget{Jparents}{\\
$J_{\rm{\rm{{parents}}}$ & 256 & This is the number of parents of each dataset element. }\hypertarget{Jcacherounds}{\\
$J_{\rm{{cacherounds}}$ & 3 & This is the number of rounds in cache production. }\hypertarget{Jaccesses}{\\
$J_{\rm{{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop. }\\
\bottomrule
\end{tabu}
%\end{tabularx}

\subsection{Size of dataset and cache}
The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
\begin{equation}
E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{\rm{epoch}}}}\right\rfloor
\end{equation}
The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\rm{{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_\rm{{cachegrowth}}$} bytes, where the growth of each occurs every epoch. In order to avoid regularity leading to cyclic behavior, the size must be a prime number. Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_\rm{{hashbytes}}$} for the cache.\hypertarget{dsize}{
Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
\begin{equation}
 d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\rm{{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\rm{{datasetgrowth}} \cdot E_\rm{{epoch}} - \hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}, J_\rm{{mixbytes}})
\end{equation}
The size of the cache, $c_{size}$, is calculated using
\begin{equation}
 c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_\rm{{cacheinit}}} + \hyperlink\rm{{cachegrowth}}{J_\rm{{cachegrowth}}} \cdot E_\rm{{epoch}} - \hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}, J_\rm{{hashbytes}})
\end{equation}
\begin{equation}
 E_{prime}(x, y) = \begin{cases}
x & \text{if} \quad x / y \in \mathbb{P} \\
E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
\end{cases}
\end{equation}
\subsection{Dataset generation}
In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $s \in \mathbb{B}_{32}$.
\subsubsection{Seed hash}
The seed hash is different for every epoch. For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash:
\begin{equation}
 s = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
\end{equation}
\begin{equation}
 C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_\rm{{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_\rm{{epoch}}})) & \text{otherwise}
\end{cases}
\end{equation}
With $\mathbf{0}_{32}$ being 32 bytes of zeros.

\subsubsection{Cache}
The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_\rm{{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}. The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.

We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
\begin{equation}
 \mathbf{c}_{i} = \begin{cases}
\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
\end{cases}
\end{equation}
Therefore $ \mathbf{c'}$ can be defined as
\begin{equation}
 \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
\end{equation}
\begin{equation}
 n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
\end{equation}
The cache is calculated by performing \hyperlink{Jcacherounds}{$J_\rm{{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
\begin{equation}
 \mathbf{c} = E_\rm{{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_\rm{{cacherounds}}})
\end{equation}
\begin{equation}
 E_\rm{{cacherounds}}(\mathbf{x}, y) = \begin{cases}
\mathbf{x} & \text{if} \quad y = 0 \quad  \\
E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
E_\rm{{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
\end{cases}
\end{equation}
Where a single round modifies each subset of the cache as follows:
\begin{equation}
 E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
\end{equation}
\hypertarget{Ecacherounds}{}
\begin{multline}
  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
\end{multline}

\subsubsection{Full dataset calculation} \label{dataset}
Essentially, we combine data from $J_\rm{{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset. The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_\rm{{hashbytes}}$} bytes in size:
\begin{equation}
 \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
\end{equation}
In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
\begin{equation}
 E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
\end{equation}
The single item of the dataset can now be calculated as:
\begin{equation}
 E_{datasetitem}(\mathbf{c}, i) = E_\rm{{parents}}(\mathbf{c}, i, -1, \varnothing)
\end{equation}
\begin{equation}
  E_\rm{{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
E_\rm{{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_\rm{{parents}} -2 \\
E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
\end{cases}
\end{equation}
\begin{equation}
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_\rm{{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\rm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
\end{cases}
\end{equation}

\subsection{Proof-of-work function}
Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix. \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.

@@

The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
\begin{equation}
 \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
\end{equation}
With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
\begin{equation}
 \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_\rm{{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
\end{equation}

>>>>

%\usepackage{longtable}
%https://stackoverflow.com/questions/1918366/automated-line-breaks-in-latex-tables/1918465#1918465
%tabularx seems better, but we can include this if we need tables to break across pages automatically, as well as tabu and longtabu

@@

%was using tabularx, but found tabu which seems more flexible, and generally better.
\usepackage{tabu}

@@

\begin{tabularx}{\textwidth}{rlrrX}
\midrule
0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}. Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$. As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. Otherwise it is exactly equivalent to {\small CALL} except: \\
&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{s}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_\mathbf{p}, 0, \hyperlink{I v}{I_\mathbf{v}}, \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathbf{v}} \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$. This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
\midrule
0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{SELFDESTRUCT}{}\\
\midrule
0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_\mathbf{a} \}$ \\
&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} + \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ \\
&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{b} \equiv 0$ \\
&&&& $A'_{r} \equiv A_{r} + \begin{cases}
\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
0 & \text{otherwise} 
\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{selfdestruct}} + \begin{cases}
\hyperlink{Gnewaccount}{G_{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
0 & \text{otherwise}
\end{cases}$ \\
\bottomrule
\end{tabularx}

@@

\section{Ethash}\label{app:ethash}
\subsection{Definitions}
We employ the following definitions:

\begin{tabularx}{\textwidth}{lrX}
%\begin{tabu}{X[,l]X[,r]X[,l]}

\toprule
Name & Value & Description\hypertarget{Jwordbytes}{}\\
\midrule 
$J_{\rm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{\\
$J_{\rm{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis.}\hypertarget{Jdatasetgrowth}{\\
$J_{\rm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch.}\hypertarget{Jcacheinit}{\\
$J_{\rm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
$J_{\rm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
$J_{\rm{epoch}}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
$J_{\rm{mixbytes}}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
$J_{\rm{hashbytes}}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
$J_{\rm{parents}}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
$J_{\rm{cacherounds}}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
$J_{\rm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
\bottomrule
%\end{tabu}
\end{tabularx}

\subsection{Size of dataset and cache}
The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
\begin{equation}
E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{\rm{epoch}}}}\right\rfloor
\end{equation}
The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\rm{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\rm{cachegrowth}}$} bytes, where the growth of each occurs every epoch. In order to avoid regularity leading to cyclic behavior, the size must be a prime number. Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\rm{hashbytes}}$} for the cache.\hypertarget{dsize}{
Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
\begin{equation}
 d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\rm{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\rm{datasetgrowth}} \cdot E_{\rm{epoch}} - \hyperlink{Jmixbytes}{J_\{rm{mixbytes}}}, J_{\rm{mixbytes}})
\end{equation}
The size of the cache, $c_{size}$, is calculated using
\begin{equation}
 c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\rm{cacheinit}}} + \hyperlink{\rm{cachegrowth}}{J_{\rm{cachegrowth}}} \cdot E_{\rm{epoch}} - \hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}, J_{\rm{hashbytes}})
\end{equation}
\begin{equation}
 E_{prime}(x, y) = \begin{cases}
x & \text{if} \quad x / y \in \mathbb{P} \\
E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
\end{cases}
\end{equation}
\subsection{Dataset generation}
In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $s \in \mathbb{B}_{32}$.
\subsubsection{Seed hash}
The seed hash is different for every epoch. For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash:
\begin{equation}
 s = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
\end{equation}
\begin{equation}
 C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\rm{epoch}}})) & \text{otherwise}
\end{cases}
\end{equation}
With $\mathbf{0}_{32}$ being 32 bytes of zeros.

\subsubsection{Cache}
The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\rm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}. The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.

We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
\begin{equation}
 \mathbf{c}_{i} = \begin{cases}
\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
\end{cases}
\end{equation}
Therefore $ \mathbf{c'}$ can be defined as
\begin{equation}
 \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
\end{equation}
\begin{equation}
 n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
\end{equation}
The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\rm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
\begin{equation}
 \mathbf{c} = E_{\rm{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{\rm{cacherounds}}})
\end{equation}
\begin{equation}
 E_{\rm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
\mathbf{x} & \text{if} \quad y = 0 \quad  \\
E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
E_{\rm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
\end{cases}
\end{equation}
Where a single round modifies each subset of the cache as follows:
\begin{equation}
 E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
\end{equation}
\hypertarget{Ecacherounds}{}
\begin{multline}
  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
\end{multline}

\subsubsection{Full dataset calculation} \label{dataset}
Essentially, we combine data from $J_{\rm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset. The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{\rm{hashbytes}}$} bytes in size:
\begin{equation}
 \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
\end{equation}
In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
\begin{equation}
 E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
\end{equation}
The single item of the dataset can now be calculated as:
\begin{equation}
 E_{datasetitem}(\mathbf{c}, i) = E_{\rm{parents}}(\mathbf{c}, i, -1, \varnothing)
\end{equation}
\begin{equation}
  E_{\rm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
E_{\rm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\rm{parents}} -2 \\
E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
\end{cases}
\end{equation}
\begin{equation}
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\rm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\rm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
\end{cases}
\end{equation}

\subsection{Proof-of-work function}
Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix. \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.

@@

The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
\begin{equation}
 \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
\end{equation}
With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
\begin{equation}
 \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\rm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
\end{equation}
